Drivers of spatial synchrony in giant kelp populations

### 2021-05-13 paper/SupMat.Rnw
changes to supmat:

%packages
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage[left=2cm,top=2cm,right=2cm,bottom=2cm,head=.5cm,foot=.5cm]{geometry}
\usepackage{url}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{subfig}
\usepackage{float}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{xr}
\usepackage{authblk}
%\usepackage{accents}

@@

\noindent for species $i=1,2$. Here, $N_i(t)$ is the population density of species $i$ at time $t$, and $N = N_1(t)+N_2(t)$ is fixed through time. The parameter, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundiity of specise $i$ at time $t$. The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles in population to the reproductive output of the two species that year. We take, for each $i$, $B_i(t)$ to be $iid$ through time. We let $B_i = \exp(b_i)$, where $(b_1, b_2)$ is some distribution such that $b_i \sim N(\mu_i, \sigma^2)$ and such that $cov(b_1, b_2) = \rho$, a fixed quantity we will specify later. However, $(b_1, b_2)$ is not necessarily a bivariate normal distribution. We will consider various distributions of $(b_1, b_2)$ with these properties, corresponding to our symmetric and asymmetric tail associations cases defined in the main text and elaborated below. We assume without loss of generality that $\mu_1 \leq \mu_2$, so that species 1 is the weaker competitor.

@@

We created three cases for $(b_1,b_2)$, the case of symmetric tail association and the cases of asymmetrical left- and right-tail association. To generate $M$ points $(b_1^{(i)}, b_2^{(i)})$, $i=1,..,M$, for the left tail association case, we first generated M points $(a_1^{(i)},a_2^{(i)})$, $i=1,...,M$, from a bivariate normal distribution $N(\begin{pmatrix} 0\\0\end{pmatrix}, \Sigma)$, where here $\Sigma = \begin{pmatrix} 1&0\\0&1\end{pmatrix}$. Then, for each index $i$, we randomly, with 50\% probability, either (A) let $\tilde b_1^{(i)} = -|a_1^{(i)}|$ and $\tilde b_2^{(i)} = -a|_1^{(i)}|$, or (B) let $\tilde b_1^{(i)} = |a_1^{(i)}|$ and $\tilde b_2^{(i)} = |a_2^{(i)}|$. We then let $b_1^{(i)} = \sigma \tilde b_1^{(i)} + \mu_1$ and $b_2^{(i)} = \sigma \tilde b_2^{(i)} + \mu_2$. Perfect association in the left tails of the resulting distribution $(b_1,b_2)$ results from the fact that, in case (A) above, both $\tilde b_1^{(i)}$ and $\tilde b_2^{(i)}$ were assigned to the same value. For speed of computation, when samples $(\tilde b_1^{(i)},\tilde b_2^{(i)})$, $i=1,...,M$ were needed for every distinct set of values of the parameters $\mu_1$, $\mu_2$, and $\sigma^2$, we used the same $(\tilde b_1^{(i)},\tilde b_2^{(i)})$, $i=1,...,M$, simply transforming these values differently (using the different valus of $\mu_1$, $\mu_2$, $\sigma$) for each set of parameter values. Right-tail associated noise was generated in an analogous way.

Noise with symmetric tail association was generated by taking a very large sample (100000) of left-tail associated noise, as described above, and then calculating the Pearson correaltion coefficient, rho. The value obtained was ... . The symmetric noise was then generated once from $N(\begin{pmatrix} 0\\0\end{pmatrix}, \Sigma)$ where $\Sigma = \begin{pmatrix} 1&\rho \\ \rho&1\end{pmatrix}$, denoting the points so generated by $(\tilde b_1^{(i)},\tilde b_2^{(i)})$ for $i=1,...,M$. We then set $b_1^{(i)} = \sigma \tilde b_1^{(i)} + \mu_1$ and $b_2^{(i)} = \sigma \tilde b_2^{(i)}+ \mu_2$.

@@

First, for a large ?artigen? $M$, carry out the following steps once:

@@

We also considered the probability $\Prob[r_1 >0]$, which is, explicitly, the probability that environmental variables that permit invasion. We have
\begin{align}
\Prob[r_1>0] &= \Prob [1-\delta+\delta \exp(b_1-b_2)] > 1] \\
&= \Prob[\exp (b_1-b_2)>1] \\
&= \Prob[b_1>b_2],
\end{align}
?? should the above have $\ln$?
\noindent so this was computed for each considered parameter set of $\delta, \mu_1, \mu_2, \sigma$, and for each of our three tail association cases.

@@

Equations \ref{sym_rbar1} - \ref{sym_rbar2sharp} obviously depend only on $\mu_1-\mu_2$, and not independently on $\mu_1$ and $\mu_2$, since only $\mu_1-\mu_2$ appears in these expressions if $\mu_1$ or $\mu_2$ appear at all. Equation \ref{sym_rbar1} is $\bar r_1$ in the asymmetric? tail association cases. But letting $\beta_1 = b_1 + \eta$ and $\beta_2 = b_2 + \eta$ and substituting into \ref{rbar1_E}, we get

@@

To see that our various metrics are the same for out left- and right-tail associated noise processes, let $(b_1,b_2)$ denote our left-tail associated random variable with parameters $\mu_1,\mu_2,\sigma$. Then define $\beta_1 = -b_2 + \mu_1 +\mu_2$, $\beta_2=-b_1+\mu_1+\mu_2$. It is easy to see that $\EX(\beta_1)=\mu_1$, $\EX(\beta_2)=\mu_2$, $\sd(\beta_1)=\sd(\beta_2)=\sigma$, so $\beta_1$ and $\beta_2$ are normally distributed. From there it is easy to see that $\beta_1$ and $beta_2$ are distributed in the same wat as our right-tail associated noise with parameter $\mu_1,\mu_2$, and $\sigma$. Making the substitutions $b_2 = -\beta_1 + \mu_1 + \mu_2$ and $b_1 = -\beta_2 + \mu_1+\mu_2$ in ref{r1distribution}, we get

>>>>

%packages
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage[left=2cm,top=2cm,right=2cm,bottom=2cm,head=.5cm,foot=.5cm]{geometry}
\usepackage{url}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{subfig}
\usepackage{float}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{xr}
%\usepackage{authblk}
%\usepackage{accents}

@@

\noindent for species $i=1,2$. Here, $N_i(t)$ is the population density of species $i$ at time $t$, and $N = N_1(t)+N_2(t)$ is fixed through time. The parameter, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundiity of specise $i$ at time $t$. The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles in population to the reproductive output of the two species that year. We take, for each $i$, $B_i(t)$ to be $iid$ through time. We let $B_i = \exp(b_i)$, where $(b_1, b_2)$ is some distribution such that $b_i \sim N(\mu_i, \sigma^2)$ and such that $\CovX(b_1, b_2) = \rho$, a fixed quantity we will specify later. However, $(b_1, b_2)$ is not necessarily a bivariate normal distribution. We will consider various distributions of $(b_1, b_2)$ with these properties, corresponding to our symmetric and asymmetric tail associations cases defined in the main text and elaborated below. We assume without loss of generality that $\mu_1 \leq \mu_2$, so that species 1 is the weaker competitor.

@@

We created three cases for $(b_1,b_2)$, the case of symmetric tail association and the cases of asymmetrical left- and right-tail association. To generate $M$ points $(b_1^{(i)}, b_2^{(i)})$, $i=1,..,M$, for the left tail association case, we first generated M points $(a_1^{(i)},a_2^{(i)})$, $i=1,...,M$, from a bivariate normal distribution $N([0,0], \Sigma)$, where here $\Sigma = \begin{pmatrix} 1&0\\0&1\end{pmatrix}$. Then, for each index $i$, we randomly, with 50\% probability, either (A) let $\tilde b_1^{(i)} = -|a_1^{(i)}|$ and $\tilde b_2^{(i)} = -a|_1^{(i)}|$, or (B) let $\tilde b_1^{(i)} = |a_1^{(i)}|$ and $\tilde b_2^{(i)} = |a_2^{(i)}|$. We then let $b_1^{(i)} = \sigma \tilde b_1^{(i)} + \mu_1$ and $b_2^{(i)} = \sigma \tilde b_2^{(i)} + \mu_2$. Perfect association in the left tails of the resulting distribution $(b_1,b_2)$ results from the fact that, in case (A) above, both $\tilde b_1^{(i)}$ and $\tilde b_2^{(i)}$ were assigned to the same value. For speed of computation, when samples $(\tilde b_1^{(i)},\tilde b_2^{(i)})$, $i=1,...,M$ were needed for every distinct set of values of the parameters $\mu_1$, $\mu_2$, and $\sigma^2$, we used the same $(\tilde b_1^{(i)},\tilde b_2^{(i)})$, $i=1,...,M$, simply transforming these values differently (using the different valus of $\mu_1$, $\mu_2$, $\sigma$) for each set of parameter values. Right-tail associated noise was generated in an analogous way.

Noise with symmetric tail association was generated by taking a very large sample (100000) of left-tail associated noise, as described above, and then calculating the Pearson correaltion coefficient, rho. The value obtained was ... . The symmetric noise was then generated once from $N([0,0], \Sigma)$ where $\Sigma = \begin{pmatrix} 1&\rho \\ \rho&1\end{pmatrix}$, denoting the points so generated by $(\tilde b_1^{(i)},\tilde b_2^{(i)})$ for $i=1,...,M$. We then set $b_1^{(i)} = \sigma \tilde b_1^{(i)} + \mu_1$ and $b_2^{(i)} = \sigma \tilde b_2^{(i)}+ \mu_2$.

@@

First, for a large integer $M$, carry out the following steps once:

@@

We also considered the probability $\Prob[r_1 >0]$, which is, explicitly, the probability that environmental variables that permit invasion. We have
\begin{align}
\Prob[r_1>0] &= \Prob [1-\delta+\delta \exp(b_1-b_2)] > 1] \\
&= \Prob[\exp (b_1-b_2)>1] \\
&= \Prob[b_1>b_2],
\end{align}

\noindent so this was computed for each considered parameter set of $\delta, \mu_1, \mu_2, \sigma$, and for each of our three tail association cases.

@@

Equations \ref{sym_rbar1} - \ref{sym_rbar2sharp} obviously depend only on $\mu_1-\mu_2$, and not independently on $\mu_1$ and $\mu_2$, since only $\mu_1-\mu_2$ appears in these expressions if $\mu_1$ or $\mu_2$ appear at all. Equation \ref{rbar1_E} is $\bar r_1$ in the asymmetric tail association cases. But letting $\beta_1 = b_1 + \eta$ and $\beta_2 = b_2 + \eta$ and substituting into \ref{rbar1_E}, we get

@@

To see that our various metrics are the same for out left- and right-tail associated noise processes, let $(b_1,b_2)$ denote our left-tail associated random variable with parameters $\mu_1,\mu_2,\sigma$. Then define $\beta_1 = -b_2 + \mu_1 +\mu_2$, $\beta_2=-b_1+\mu_1+\mu_2$. It is easy to see that $\EX(\beta_1)=\mu_1$, $\EX(\beta_2)=\mu_2$, $\sd(\beta_1)=\sd(\beta_2)=\sigma$, so $\beta_1$ and $\beta_2$ are normally distributed. From there it is easy to see that $\beta_1$ and $beta_2$ are distributed in the same wat as our right-tail associated noise with parameter $\mu_1,\mu_2$, and $\sigma$. Making the substitutions $b_2 = -\beta_1 + \mu_1 + \mu_2$ and $b_1 = -\beta_2 + \mu_1+\mu_2$ in \ref{r1distribution}, we get

### 2021-05-13 paper/SupMat.tex
changes to supmat:

\noindent for species $i=1,2$. Here, $N_i(t)$ is the population density of species $i$ at time $t$, and $N = N_1(t)+N_2(t)$ is fixed through time. The parameter, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundiity of specise $i$ at time $t$. The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles in population to the reproductive output of the two species that year. We take, for each $i$, $B_i(t)$ to be $iid$ through time. We let $B_i = exp(b_i)$, where $(b_1, b_2)$ is some distribution such that $b_i \sim N(\mu_i, \sigma^2)$ and such that $cov(b_1, b_2) = \rho$, a fixed quantity we will specify later. However, $(b_1, b_2)$ is not necessarily a bivariate normal distribution. We will consider various distributions of $(b_1, b_2)$ with these properties, corresponding to our symmetric and asymmetric tail associations cases defined in the main text and elaborated below. We assume without loss of generality that $\mu_1 \leq \mu_2$, so that species 1 is the weaker competitor.

@@

Equations \ref{sym_rbar1} - \ref{sym_rbar2sharp} obviously depend only on $\mu_1-\mu_2$, and not independently on $\mu_1$ and $\mu_2$, ?some? only $\mu_1-\mu_2$ appears in these expressions if $\mu_1$ or $\mu_2$ appear at all. Equation \ref{sym_rbar1} is $\bar r_1$ in the asymmetric? tail association cases. But letting $\beta_1 = b_1 + \eta$ and $\beta_2 = b_2 + \eta$ and substituting into \ref{sym_rbar1}, we get

@@

To see that our various metrics are the same for out left- and right-tail associated noise ?parameters?, let $(b_1,b_2)$ denote our left-tail associated random variable with parameters $\mu_1,\mu_2,\sigma$. Then define $\beta_1 = -b_2 + \mu_1 +\mu_2$, $\beta_2=-b_1+\mu_1+\mu_2$. It is easy to see that $\EX(\beta_1)=\mu_1$, $\EX(\beta_2)=\mu_2$, $\sd(\beta_1)=\sd(\beta_2)=\sigma$, so $\beta_1$ and $\beta_2$ are normally distributed. From there it is easy to see that $\beta_1$ and $beta_2$ are distributed in the same wat as our right-tail associated noise with parameter $\mu_1,\mu_2$, and $\sigma$. Making the substitutions $b_2 = -\beta_1 + \mu_1 + \mu_2$ and $b_1 = -\beta_2 + \mu_1+\mu_2$ in ref{r1distribution}, we get

>>>>

\noindent for species $i=1,2$. Here, $N_i(t)$ is the population density of species $i$ at time $t$, and $N = N_1(t)+N_2(t)$ is fixed through time. The parameter, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundiity of specise $i$ at time $t$. The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles in population to the reproductive output of the two species that year. We take, for each $i$, $B_i(t)$ to be $iid$ through time. We let $B_i = \exp(b_i)$, where $(b_1, b_2)$ is some distribution such that $b_i \sim N(\mu_i, \sigma^2)$ and such that $\CovX(b_1, b_2) = \rho$, a fixed quantity we will specify later. However, $(b_1, b_2)$ is not necessarily a bivariate normal distribution. We will consider various distributions of $(b_1, b_2)$ with these properties, corresponding to our symmetric and asymmetric tail associations cases defined in the main text and elaborated below. We assume without loss of generality that $\mu_1 \leq \mu_2$, so that species 1 is the weaker competitor.

@@

Equations \ref{sym_rbar1} - \ref{sym_rbar2sharp} obviously depend only on $\mu_1-\mu_2$, and not independently on $\mu_1$ and $\mu_2$, since only $\mu_1-\mu_2$ appears in these expressions if $\mu_1$ or $\mu_2$ appear at all. Equation \ref{rbar1_E} is $\bar r_1$ in the asymmetric tail association cases. But letting $\beta_1 = b_1 + \eta$ and $\beta_2 = b_2 + \eta$ and substituting into \ref{rbar1_E}, we get

@@

To see that our various metrics are the same for out left- and right-tail associated noise processes, let $(b_1,b_2)$ denote our left-tail associated random variable with parameters $\mu_1,\mu_2,\sigma$. Then define $\beta_1 = -b_2 + \mu_1 +\mu_2$, $\beta_2=-b_1+\mu_1+\mu_2$. It is easy to see that $\EX(\beta_1)=\mu_1$, $\EX(\beta_2)=\mu_2$, $\sd(\beta_1)=\sd(\beta_2)=\sigma$, so $\beta_1$ and $\beta_2$ are normally distributed. From there it is easy to see that $\beta_1$ and $beta_2$ are distributed in the same wat as our right-tail associated noise with parameter $\mu_1,\mu_2$, and $\sigma$. Making the substitutions $b_2 = -\beta_1 + \mu_1 + \mu_2$ and $b_1 = -\beta_2 + \mu_1+\mu_2$ in \ref{r1distribution}, we get
